---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

requireNamespace("hydroGOF", quietly = TRUE)
requireNamespace("bench", quietly = TRUE)
```

# tidyhydro

<!-- badges: start -->
[![R-CMD-check](https://github.com/atsyplenkov/tidyhydro/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/atsyplenkov/tidyhydro/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/tidyhydro)](https://CRAN.R-project.org/package=tidyhydro)
![GitHub R package version](https://img.shields.io/github/r-package/v/atsyplenkov/tidyhydro?label=github)
![GitHub last commit](https://img.shields.io/github/last-commit/atsyplenkov/tidyhydro)
<!-- badges: end -->

The `tidyhydro` package provides a set of commonly used metrics in hydrology (such as NSE, KGE, pBIAS) for use within a [`tidymodels`](https://www.tidymodels.org/) infrastructure. Originally inspired by the [`yardstick`](https://github.com/tidymodels/yardstick/tree/main)and [`hydroGOF`](https://github.com/hzambran/hydroGOF) packages, this library is mainly written in C++ and provides a very quick estimation of desired goodness-of-fit criteria.

## Installation

You can install the development version of `tidyhydro` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("atsyplenkov/tidyhydro")
```

## Example

```{r example}
library(yardstick)
library(tidyhydro)
data(solubility_test)

nse(solubility_test, solubility, prediction)
```

## Benchmarking

```{r benchmarking}
bench::mark(
  tidyhydro = tidyhydro::nse_vec(
    solubility_test$solubility,
    solubility_test$prediction
  ),
  hydroGOF = hydroGOF::NSE(
    sim = solubility_test$prediction,
    obs = solubility_test$solubility
  ),
  check = TRUE,
  relative = TRUE
)

```