---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

requireNamespace("hydroGOF", quietly = TRUE)
requireNamespace("bench", quietly = TRUE)
```

# tidyhydro

<!-- badges: start -->
<p align="center">
    <a href="https://github.com/atsyplenkov/tidyhydro/releases">
        <img src="https://img.shields.io/github/v/release/atsyplenkov/tidyhydro?style=flat&labelColor=1C2C2E&color=198ce7&logo=GitHub&logoColor=white"></a>
    <!-- <a href="https://cran.r-project.org/package=tidyhydro">
        <img src="https://img.shields.io/cran/v/tidyhydro?style=flat&labelColor=1C2C2E&color=198ce7&logo=R&logoColor=white"></a> -->
    <a href="https://app.codecov.io/gh/atsyplenkov/tidyhydro">
        <img src="https://img.shields.io/codecov/c/gh/atsyplenkov/tidyhydro?style=flat&labelColor=1C2C2E&color=256bc0&logo=Codecov&logoColor=white"></a>
    <a href="https://github.com/atsyplenkov/tidyhydro/actions/workflows/check-r-pkg.yaml">
        <img src="https://img.shields.io/github/actions/workflow/status/atsyplenkov/tidyhydro/check-r-pkg.yaml?style=flat&labelColor=1C2C2E&color=256bc0&logo=GitHub%20Actions&logoColor=white"></a>
</p>
<!-- badges: end -->

The `tidyhydro` package provides a set of commonly used metrics in hydrology (such as _NSE_, _KGE_, _pBIAS_) for use within a [`tidymodels`](https://www.tidymodels.org/) infrastructure. Originally inspired by the [`yardstick`](https://github.com/tidymodels/yardstick/tree/main) and [`hydroGOF`](https://github.com/hzambran/hydroGOF) packages, this library is mainly written in C++ and provides a very quick estimation of desired goodness-of-fit criteria.

Additionally, you'll find here a C++ implementation of lesser-known yet powerful metrics used in reports from the United States Geological Survey (USGS) and the National Environmental Monitoring Standards (NEMS) guidelines. Examples include _PRESS_ (Prediction Error Sum of Squares), _SFE_ (Standard Factorial Error), and _MSPE_ (Model Standard Percentage Error) and others. Based on the equations from _Helsel et al._ ([2020](https://pubs.usgs.gov/publication/tm4A3)), _Rasmunsen et al._ ([2008](https://pubs.usgs.gov/tm/tm3c4/)), _Hicks et al._ ([2020](https://www.nems.org.nz/documents/suspended-sediment)) and etc. (see documentation for details).

## Example
The `tidyhydro` package follows the philosophy of [`yardstick`](https://github.com/tidymodels/yardstick/tree/main) and provides S3 class methods for vectors and data frames. For example, one can estimate `NSE` and `pBIAS` for a data frame like this:

```{r example}
library(tidyhydro)
data(avacha)

avacha

kge(avacha, obs, sim)
```

or create a [`metric_set`](https://yardstick.tidymodels.org/reference/metric_set.html) and estimate several parameters at once like this:

```{r metricset}
hydro_metrics <- yardstick::metric_set(nse, pbias)

hydro_metrics(avacha, obs, sim)
```

We do understand that sometimes one needs a qualitative interpretation of the model. Therefore, we populated some functions with a `performance` argument. When `performance = TRUE`, the metric interpretation will be returned according to Moriasi et al. ([2015](https://elibrary.asabe.org/abstract.asp?aid=46548&t=3&dabs=Y&redir=&redirType=)).

```{r interpretation}
hydro_metrics(avacha, obs, sim, performance = TRUE)
```

## Installation

You can install the development version of `tidyhydro` from [GitHub](https://github.com/atsyplenkov/tidyhydro) with:

``` r
# install.packages("pak")
pak::pak("atsyplenkov/tidyhydro")
```

## Benchmarking
Since the package uses `Rcpp` in the background, it performs slightly faster than base R and other R packages (see [benchmarks](https://atsyplenkov.github.io/tidyhydro/articles/benchmarks.html)). This is particularly noticeable with large datasets:
```{r benchmarking}
set.seed(12234)
x <- runif(10^6)
y <- runif(10^6)

nse <- function(truth, estimate, na_rm = TRUE) {
  #fmt: skip
  1 - (sum((truth - estimate)^2, na.rm = na_rm) /
        sum((truth - mean(truth, na.rm = na_rm))^2, na.rm = na_rm))
}

bench::mark(
  tidyhydro = tidyhydro::nse_vec(truth = x, estimate = y),
  hydroGOF = hydroGOF::NSE(sim = y, obs = x),
  baseR = nse(truth = x, estimate = y),
  check = TRUE,
  relative = TRUE,
  filter_gc = FALSE,
  iterations = 50L
)
```

## Code of Conduct
  
Please note that the tidyhydro project is released with a [Contributor Code of Conduct](https://atsyplenkov.github.io/tidyhydro/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

## See also
* [`hydroGOF`](https://github.com/hzambran/hydroGOF) - Goodness-of-fit functions for comparison of simulated and observed hydrological time series.
* [`yardstick`](https://github.com/tidymodels/yardstick/tree/main) - tidy methods for models performance assessment.
