---
title: "Getting started"
format:
  html:
    toc: true
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
    message: false
---

```{r}
#| label: setup
#| include: false
library(tidyhydro)
library(ggplot2)
theme_set(
  theme_minimal() +
    theme(legend.position = "bottom")
)
```

# Available metrics

In `tidyhydro` v`r packageVersion("tidyhydro")`, `r length(getNamespaceExports("tidyhydro"))/2` metrics are implemented.

| Name | Abbr. | Function calls | Reference |
|----------------|--------|--------|-------------|
| Kling-Gupta Efficiency | $KGE$ | `kge`, `kge_vec` | <span style="font-size: 0.8em;">Gupta, H.V.; Kling, H.; Yilmaz, K.K.; Martinez, G.F. (2009). *Journal of Hydrology*, 377(1–2), 80–91.</span> |
| Modified Kling-Gupta Efficiency | $KGE'$ | `kge2012`, `kge_vec` | <span style="font-size: 0.8em;">Kling, H., Fuchs, M., & Paulin, M. (2012). *Journal of Hydrology*, 424–425, 264–277.</span> |
| Nash-Sutcliffe Efficiency | $NSE$ | `nse`, `nse_vec` | <span style="font-size: 0.8em;">Nash, J. E., & Sutcliffe, J. V. (1970). *Journal of Hydrology*, 10(3), 282–290.</span> |
| Mean Squared Error | $MSE$ | `mse`, `mse_vec` | <span style="font-size: 0.8em;">Clark, M. P., Vogel, R. M., Lamontagne, J. R., Mizukami, N., Knoben, W. J. M., Tang, G., Gharari, S., Freer, J. E., Whitfield, P. H., Shook, K. R., & Papalexiou, S. M. (2021). The Abuse of Popular Performance Metrics in Hydrologic Modeling. Water Resources Research, 57(9), e2020WR029001.</span> |
| Percent BIAS | $pBIAS$ | `pbias`, `pbias_vec` | <span style="font-size: 0.8em;">Gupta, H. V., S. Sorooshian, and P. O. Yapo. (1999). Status of automatic calibration for hydrologic models: Comparison with multilevel expert calibration. J. Hydrologic Eng. 4(2): 135-143 </span> |
| PRediction Error Sum of Squares | $PRESS$ | `press`, `press_vec` | <span style="font-size: 0.8em;"> Rasmussen, P. P., Gray, J. R., Glysson, G. D. & Ziegler, A. C. Guidelines and procedures for computing time-series suspended-sediment concentrations and loads from in-stream turbidity-sensor and streamflow data. in U.S. Geological Survey Techniques and Methods book 3, chap. C4 53 (2009)</span> |
| Standard Factorial Error | $SFE$ | `sfe`, `sfe_vec` | <span style="font-size: 0.8em;"> Herschy, R.W. 1978: Accuracy. Chapter 10 In: Herschy, R.W. (ed.) Hydrometry - principles and practices. John Wiley and Sons, Chichester, 511 p.</span> |

: Metrics currently implemented in `tidyhydro` v`r packageVersion("tidyhydro")`

# `avacha` dataset

The package includes the mean daily water discharge values (`obs` in m³/s) measured at the state gauging station Avacha River — Elizovo City (site No. 2090). Alongside the measured water discharge, the mean water discharge in the last 24 hours derived from the [GloFAS-ERA5 v4.0](https://confluence.ecmwf.int/display/CEMS/GloFAS+v4.0) reanalysis is provided (`sim`).

```{r}
#| fig-cap: Avacha River - Elizovo City hydrograph
library(ggplot2)
library(tidyhydro)
data(avacha)

avacha |>
  ggplot(aes(x = date)) +
  geom_line(aes(y = obs, colour = "Measured")) +
  geom_line(aes(y = sim, colour = "Predicted")) +
  scale_colour_brewer(name = "", palette = "Set1") +
  labs(x = "", y = "Water Discharge, m³/s")
```

# Example usage

One can estimate the desired metrics using the `tidyverse` [syntax](https://style.tidyverse.org/). For example, to get the Nash-Sutcliffe Efficiency ($NSE$) or Modified Kling-Gupta Efficiency ($KGE'$) for the `avacha` dataset, one can run:

```{r}
nse(avacha, obs, sim)
kge2012(avacha, obs, sim)
```

Or using the `yardstick` helper functions, one can create a metric set, combining it with other `yardstick` metrics, such as $R^2$:

```{r}
library(yardstick)
hydro_metrics <- metric_set(kge, pbias, rsq)
hydro_metrics(avacha, obs, sim)
```

Such syntax is particularly useful when running a group analysis, for example, estimating model performance for different months:

```{r}
library(lubridate)
library(dplyr)

avacha |>
  mutate(month = month(date)) |>
  group_by(month) |>
  nse(obs, sim)
```

Alternatively, one can still use the vectorised versions of the metrics, ending with the `*_vec` suffix:

```{r}
nse_vec(truth = avacha$obs, estimate = avacha$sim)
```